---
title: "Clinic4"
author: "Pablo Perez"
date: "3/11/2018"
output: html_document
---

#Loading libraries
```{r, message=FALSE}
#source("https://bioconductor.org/biocLite.R")
#biocLite("EBImage")
#biocLite("CRImage")

library(EBImage)
library(CRImage)
```


#Image Basics and Histograms, (Image1).
```{r}
image = readImage("image1.jpg")
display(image, method="raster")
plot(image)
str(image)

#Convert image to gray values
grayImage <- EBImage::channel(image,"gray")

#The resulting matrix has size 640x427, 3 times smaller than the colored image matrix with size 640x427x3
str(grayImage)
plot(grayImage)

#binary image
lenab=grayImage
lenab[lenab>0.2]=1
lenab[lenab<=0.2]=0

lenab1=grayImage
lenab1[lenab1>0.4]=1
lenab1[lenab1<=0.4]=0

lenab2=grayImage
lenab2[lenab2>0.6]=1
lenab2[lenab2<=0.6]=0

lenab3=grayImage
lenab3[lenab3>0.8]=1
lenab3[lenab3<=0.8]=0

lenab4=grayImage
lenab4[lenab4>0.9]=1
lenab4[lenab4<=0.9]=0

par(mfrow=c(3,3),mar=c(0, 0, 3, 0))
plot.new()
plot.window(xlim=c(0, 0.5), ylim=c(0, 0.5), asp=1)
rasterImage(grayImage, 0, 0, 1, 1)
title("original image")

plot.new()
plot.window(xlim=c(0, 0.5), ylim=c(0, 0.5), asp=1)
rasterImage(lenab, 0, 0, 1, 1)
title("threshold 0.2 image")

plot.new()
plot.window(xlim=c(0, 0.5), ylim=c(0, 0.5), asp=1)
rasterImage(lenab1, 0, 0, 1, 1)
title("threshold 0.4 image")

plot.new()
plot.window(xlim=c(0, 0.5), ylim=c(0, 0.5), asp=1)
rasterImage(lenab2, 0, 0, 1, 1)
title("threshold 0.6 image")

plot.new()
plot.window(xlim=c(0, 0.5), ylim=c(0, 0.5), asp=1)
rasterImage(lenab3, 0, 0, 1, 1)
title("threshold 0.8 image")

plot.new()
plot.window(xlim=c(0, 0.5), ylim=c(0, 0.5), asp=1)
rasterImage(lenab4, 0, 0, 1, 1)
title("threshold 0.9 image")

```


#Image Basics and Histograms, (own mountain image).
```{r}
image = readImage("mountain.jpg")
display(image, method="raster")
plot(image)
str(image)

#Convert image to gray values
grayImage2 <- EBImage::channel(image,"gray")

#The resulting matrix has size 400x400, 3 times smaller than the colored image matrix with size 400x400x3
str(grayImage2)
plot(grayImage2)

#binary image
lenab=grayImage2
lenab[lenab>0.2]=1
lenab[lenab<=0.2]=0

lenab1=grayImage2
lenab1[lenab1>0.4]=1
lenab1[lenab1<=0.4]=0

lenab2=grayImage2
lenab2[lenab2>0.6]=1
lenab2[lenab2<=0.6]=0

lenab3=grayImage2
lenab3[lenab3>0.8]=1
lenab3[lenab3<=0.8]=0

lenab4=grayImage2
lenab4[lenab4>0.9]=1
lenab4[lenab4<=0.9]=0

par(mfrow=c(3,3),mar=c(0, 0, 3, 0))
plot.new()
plot.window(xlim=c(0, 0.5), ylim=c(0, 0.5), asp=1)
rasterImage(grayImage2, 0, 0, 1, 1)
title("original image")

plot.new()
plot.window(xlim=c(0, 0.5), ylim=c(0, 0.5), asp=1)
rasterImage(lenab, 0, 0, 1, 1)
title("threshold 0.2 image")

plot.new()
plot.window(xlim=c(0, 0.5), ylim=c(0, 0.5), asp=1)
rasterImage(lenab1, 0, 0, 1, 1)
title("threshold 0.4 image")

plot.new()
plot.window(xlim=c(0, 0.5), ylim=c(0, 0.5), asp=1)
rasterImage(lenab2, 0, 0, 1, 1)
title("threshold 0.6 image")

plot.new()
plot.window(xlim=c(0, 0.5), ylim=c(0, 0.5), asp=1)
rasterImage(lenab3, 0, 0, 1, 1)
title("threshold 0.8 image")

plot.new()
plot.window(xlim=c(0, 0.5), ylim=c(0, 0.5), asp=1)
rasterImage(lenab4, 0, 0, 1, 1)
title("threshold 0.9 image")
```

#Functions
```{r}
#Contrast sketching
contrast <- function(contrast, a, b, y.a, y.b){
  
  alpha <- 0.2
  beta <- 2
  gamma <- 1
  
  contrast[contrast<a]=a*contrast
  contrast[contrast<b && contrast>=a]=(beta*(contrast-a) + y.a)
  contrast[contrast>=b]=(gamma*(contrast-b) +y.b)
  
  return(display(normalize(contrast), method = "raster"))

}
image1=floor(grayImage * 255)
image2=floor(grayImage2 * 255)
contrast(image1, 50,150,30,200)
contrast(image2, 50,150,30,200)
#Clipping
clipping <- function(clipping, a, b){
  beta <- 2
  
  clipping[clipping<a]=0
  clipping[clipping<b && clipping>=a]=beta*(clipping-a)
  clipping[clipping>=b]=beta*(b-a)
  
  return(display(normalize(clipping), method = "raster"))

}
image1=floor(grayImage * 255)
image2=floor(grayImage2 * 255)
clipping(image1,50,150)
clipping(image2,50,150)


#Range compression
range.compression <- function(comp, c){
 
  comp = c*log10(1+comp)
  return(display(normalize(comp), method="raster"))
}

image1=floor(grayImage * 255)
image2=floor(grayImage2 * 255)
range.compression(image1,100)
range.compression(image2,100)

histimg=floor(grayImage * 255)
histimg2=floor(grayImage2 * 255)

#Comparison between gray scale image and its histograms with different bins (image1)
par(mfrow=c(2,2))
display(normalize(histimg),method = "raster")
hist(histimg, breaks=c(0:255),main="Histogram with 255 breaks")
hist(histimg, breaks=55,main="Histogram with 55 breaks")
hist(histimg,breaks=5,main="Histogram with 5 breaks")

#Comparison between gray scale image and its histograms with different bins (own mountain image)
par(mfrow=c(2,2))
display(normalize(histimg2),method = "raster")
hist(histimg2, breaks=c(0:255),main="Histogram with 255 breaks")
hist(histimg2, breaks=55,main="Histogram with 55 breaks")
hist(histimg2,breaks=5,main="Histogram with 5 breaks")

#Histogram equalization
histeq <- function(x) {
  # Supporse that x is a grayscale image (range 0 % 255)
  L=256
  
  M=dim(x)[1]
  N=dim(x)[2]
  h=array(0,L+1)
  
  #This part computes the histogram
  for (i in (1:L)) {
    h[i]=length(which(x==(i-1)));
  }
  
  #This part does the equalization
  y=x;
  s=sum(h);
  
  for (i in (1:L))  {
    getthemall=which(x==(i-1));    
    y[getthemall]=(sum(h[1:i])/s*(L-1));
  }
  return (floor(y))
}

equa=floor(grayImage * 255)
eq = histeq(equa)
range(eq)

#Image and histogram for the equalization of the gray image (Image1)
par(mfrow=c(2,2))
display(normalize(equa), method="raster")
display(normalize(eq), method="raster")
hist(equa, breaks=c(0:255))
hist(eq, breaks=c(0:255))

equa=floor(grayImage2 * 255)
eq = histeq(equa)
range(eq)

#Image and histogram for the equalization of the gray image (own mountain image)
par(mfrow=c(2,2))
display(normalize(equa), method="raster")
display(normalize(eq), method="raster")
hist(equa, breaks=c(0:255))
hist(eq, breaks=c(0:255))
```

#Filters and Convolution
```{r}
#Read image and convert to gray scale
image2 = readImage("image2.jpg")
grayImage2 <- EBImage::channel(image2,"gray")
plot(grayImage2)

#take row 200 and plot it (image2)
row200 <- grayImage2[,200]
plot(row200)

#vector for convolution (image2)
vector <- c(0,0,0,1,0,0,0,0,0)

#convolution filtering to the row (image2)
filter <- filter(row200,vector, method = "convolution")
convolved <- EBImage::channel(filter,"gray")
plot(convolved)
grayImage2[,200] <- filter
plot(grayImage2)

#Same procedure for own image
mountain = readImage("mountain.jpg")
grayImage2 <- EBImage::channel(mountain,"gray")
plot(grayImage2)

#take row 200 and plot it (own mountain image)
row200 <- grayImage2[,200]
plot(row200)

#vector for convolution
vector <- c(0,0,0,1,0,0,0,0,0)

#convolution filtering to the row
filter <- filter(row200,vector, method = "convolution")
convolved <- EBImage::channel(filter,"gray")
plot(convolved)
grayImage2[,200] <- filter
plot(grayImage2)
```
As we can see if we take a really close look at the original image and then the image after convolution, the row 200 gets shifted to the left due to the vector used.

#2D and 1D filtering
```{r}
#Original gray image
image2 = readImage("image2.jpg")
grayImage2 <- EBImage::channel(image2,"gray")
plot(grayImage2)

#9x9 filter 
filter9 =
c(c(0.1, 0, 0, 0, 0, 0, 0, 0, 0),
  c(0, 0.1, 0, 0, 0, 0, 0, 0, 0),
  c(0, 0, 0.1, 0, 0, 0, 0, 0, 0),
  c(0, 0, 0, 0.1, 0, 0, 0, 0, 0),
  c(0, 0, 0, 0, 0.1, 0, 0, 0, 0),
  c(0, 0, 0, 0, 0, 0.1, 0, 0, 0),
  c(0, 0, 0, 0, 0, 0, 0.1, 0, 0),
  c(0, 0, 0, 0, 0, 0, 0, 0.1, 0),
  c(0, 0, 0, 0, 0, 0, 0, 0, 0.2))

#convolving the original image with the 9x9 filter
convolved2 <- filter(grayImage2,filter9, method = "convolution")
final1 <- as.Image(convolved2)
plot(final1)

#convolving the original image with the 1D size 10 filter
filter1<-c(0.1,0.1,0.1,0.1,0.1,0.1,0.1,0.1,0.1,0.1)
convolved3 <- filter(grayImage2,filter1, method = "convolution")
final2 <- as.Image(convolved3)
plot(final2)

transp1 = t(filter9)
transp2 = t(filter1)

#convolving the first filtered image with the transpose of the filter
trans1 <- filter(final1,transp1, method = "convolution")
t.final1 <- as.Image(trans1)
plot(t.final1)

#convolving the second filtered image with the transpose of the filter
trans2 <- filter(final2,transp2, method = "convolution")
t.final2 <- as.Image(trans2)
plot(t.final2)


#Same procedure for own image
#Original gray image
image2 = readImage("mountain.jpg")
grayImage2 <- EBImage::channel(image2,"gray")
plot(grayImage2)

#9x9 filter 
filter9 =
c(c(0.1, 0, 0, 0, 0, 0, 0, 0, 0),
  c(0, 0.1, 0, 0, 0, 0, 0, 0, 0),
  c(0, 0, 0.1, 0, 0, 0, 0, 0, 0),
  c(0, 0, 0, 0.1, 0, 0, 0, 0, 0),
  c(0, 0, 0, 0, 0.1, 0, 0, 0, 0),
  c(0, 0, 0, 0, 0, 0.1, 0, 0, 0),
  c(0, 0, 0, 0, 0, 0, 0.1, 0, 0),
  c(0, 0, 0, 0, 0, 0, 0, 0.1, 0),
  c(0, 0, 0, 0, 0, 0, 0, 0, 0.2))

#convolving the original image with the 9x9 filter
convolved2 <- filter(grayImage2,filter9, method = "convolution")
final1 <- as.Image(convolved2)
plot(final1)

#convolving the original image with the 1D size 10 filter
filter1<-c(0.1,0.1,0.1,0.1,0.1,0.1,0.1,0.1,0.1,0.1)
convolved3 <- filter(grayImage2,filter1, method = "convolution")
final2 <- as.Image(convolved3)
plot(final2)

transp1 = t(filter9)
transp2 = t(filter1)

#convolving the first filtered image with the transpose of the filter
trans1 <- filter(final1,transp1, method = "convolution")
t.final1 <- as.Image(trans1)
plot(t.final1)

#convolving the second filtered image with the transpose of the filter
trans2 <- filter(final2,transp2, method = "convolution")
t.final2 <- as.Image(trans2)
plot(t.final2)

```
When convolving an image with its filter each row gets shifted independently which makes the photo blurry, if the filter gets transpose, then you might shift the columns instead of the rows or shift the rows in different directions. Therefore a different type of blurriness occurs, normally causing the image to appear more blurry.

#Highpass filtering
```{r}

#Original gray image
image2 = readImage("image2.jpg")
grayImage2 <- EBImage::channel(image2,"gray")
plot(grayImage2)

#the high pass (hp) filter
hp=c(c(0, -1, 0),c(-1, 4, -1),c(0, -1, 0))

#Convolve the image with the hp filter and then normalize the results
highpass <- filter(grayImage2,hp,method="convolution")
hpImage <- as.Image(normalize(highpass))
plot(hpImage)

#Same procedure for own image
#Original gray image
image2 = readImage("mountain.jpg")
grayImage2 <- EBImage::channel(image2,"gray")
plot(grayImage2)

#the high pass (hp) filter
hp=c(c(0, -1, 0),c(-1, 4, -1),c(0, -1, 0))

#Convolve the image with the hp filter and then normalize the results
highpass <- filter(grayImage2,hp,method="convolution")
hpImage <- as.Image(normalize(highpass))
plot(hpImage)

```

#Edge Detection
```{r}
library(imager)
#Loading Lena image
image3 = readImage("image3.png")
image3= as.cimg(image3)
plot(image3)

#Gray scale image
grayImage3 <- grayscale(image3)
plot(grayImage3)

gradient <- imgradient(grayImage3,"xy")

dx <- imgradient(grayImage3,"x")
dy <- imgradient(grayImage3,"y")
magnitude <- sqrt(dx^2+dy^2)
plot(magnitude,main="Gradient magnitude for grayscale")

lenab=grayImage3
lenab[lenab>0.05]=1
lenab[lenab<=0.05]=0

dx <- imgradient(lenab,"x")
dy <- imgradient(lenab,"y")
magnitude <- sqrt(dx^2+dy^2)
plot(magnitude,main="Threshold 0.05")

lenab1=grayImage3
lenab1[lenab1>0.25]=1
lenab1[lenab1<=0.25]=0

dx <- imgradient(lenab1,"x")
dy <- imgradient(lenab1,"y")
magnitude <- sqrt(dx^2+dy^2)
plot(magnitude,main="Threshold 0.25")

lenab2=grayImage3
lenab2[lenab2>0.5]=1
lenab2[lenab2<=0.5]=0
dx <- imgradient(lenab2,"x")
dy <- imgradient(lenab2,"y")
magnitude <- sqrt(dx^2+dy^2)
plot(magnitude,main="Threshold 0.5")

lenab3=grayImage3
lenab3[lenab3>0.75]=1
lenab3[lenab3<=0.75]=0
dx <- imgradient(lenab3,"x")
dy <- imgradient(lenab3,"y")
magnitude <- sqrt(dx^2+dy^2)
plot(magnitude,main="Threshold 0.75")

lenab4=grayImage3
lenab4[lenab4>0.999]=1
lenab4[lenab4<=0.999]=0
dx <- imgradient(lenab4,"x")
dy <- imgradient(lenab4,"y")
magnitude <- sqrt(dx^2+dy^2)
plot(magnitude,main="Threhsold 1")

#Loading my own image
image3 = readImage("mountain.jpg")
image3= as.cimg(image3)
plot(image3)

#Gray scale image
grayImage3 <- grayscale(image3)
plot(grayImage3)

gradient <- imgradient(grayImage3,"xy")

dx <- imgradient(grayImage3,"x")
dy <- imgradient(grayImage3,"y")
magnitude <- sqrt(dx^2+dy^2)
plot(magnitude,main="Gradient magnitude for grayscale")

lenab=grayImage3
lenab[lenab>0.05]=1
lenab[lenab<=0.05]=0

dx <- imgradient(lenab,"x")
dy <- imgradient(lenab,"y")
magnitude <- sqrt(dx^2+dy^2)
plot(magnitude,main="Threshold 0.05")

lenab1=grayImage3
lenab1[lenab1>0.25]=1
lenab1[lenab1<=0.25]=0

dx <- imgradient(lenab1,"x")
dy <- imgradient(lenab1,"y")
magnitude <- sqrt(dx^2+dy^2)
plot(magnitude,main="Threshold 0.25")

lenab2=grayImage3
lenab2[lenab2>0.5]=1
lenab2[lenab2<=0.5]=0
dx <- imgradient(lenab2,"x")
dy <- imgradient(lenab2,"y")
magnitude <- sqrt(dx^2+dy^2)
plot(magnitude,main="Threshold 0.5")

lenab3=grayImage3
lenab3[lenab3>0.75]=1
lenab3[lenab3<=0.75]=0
dx <- imgradient(lenab3,"x")
dy <- imgradient(lenab3,"y")
magnitude <- sqrt(dx^2+dy^2)
plot(magnitude,main="Threshold 0.75")

lenab4=grayImage3
lenab4[lenab4>0.999]=1
lenab4[lenab4<=0.999]=0
dx <- imgradient(lenab4,"x")
dy <- imgradient(lenab4,"y")
magnitude <- sqrt(dx^2+dy^2)
plot(magnitude,main="Threhsold 1")

apply <- function(image, func = c("prewitt", "sobel"), orientation = c("horizontal", "vertical", "both")){
  
  prewittX <- c(c(-1,0,1),c(-1,0,1),c(-1,0,1)) 
  prewittY <- c(c(-1,-1,-1),c(0,0,0),c(1,1,1)) 
  
  sobelX <- c(c(-1,0,1),c(-2,0,2),c(-1,0,1))
  sobelY <- c(c(-1,-2,-1),c(0,0,0),c(1,2,1))
  
  
  if (func == "prewitt"){
    x <- prewittX
    y <- prewittY
  } else if(func == "sobel"){
    x <- sobelX
    y <- sobelY
  }
  

  if(orientation == "horizontal"){
    return(plot(as.Image(filter(image, x, method="convolution"))))
  } else if (orientation == "vertical"){
    return(plot(as.Image(filter(image, y, method="convolution"))))
  } else if(orientation=="both"){
    image= as.cimg(image)
    x <- filter(image, x)
    y <- filter(image, y)
    x= as.cimg(x)
    y= as.cimg(y)
    dx <- imgradient(x,"x")
    dy <- imgradient(y,"y")
    magnitude <- sqrt(dx^2+dy^2)
    return( plot(as.Image(magnitude)))
  }
}
image3 = readImage("image3.png")
grayImage3 = EBImage::channel(image3,"gray")
apply(grayImage3,func = "prewitt", orientation = "vertical")

image2 = readImage("mountain.jpg")
grayImage2 <- EBImage::channel(image2,"gray")
apply(grayImage2,func = "sobel", orientation = "vertical")

```





