---
title: "Clinic5"
author: "Pablo Perez"
date: "3/22/2018"
output: html_document
---

```{r,warning=FALSE, message=FALSE}
library(TSA)
library(glmnet)
library(DAAG)
library(MASS)
library(lattice)
library(doBy)
library(corrplot)
library(dplyr)
library(tidyr)
library(leaps)
library(ggplot2)
library(caret)
library(broom)
library(ptw)
library(pracma)
```

```{r}

gin <- read.csv2("gin.csv", header = TRUE, dec=",",sep=";")

gin.correl = ts(gin[,1],start=1968,freq=12)
plot(gin.correl,lw=2,col="blue")

acf(gin.correl, lag.max = 20, type = "correlation", plot = T, drop.lag.0 = F, main = "Correlogram of the gin data")

```
As we can see in the correlogram, there is a structure in the plot indicating that there is significant change in the time series. This is observed as lines fall way out of the significance region of 0.95 which is indicated by the blue doted lines.

```{r}

## Fit basic polynomial regression models to both 
tl = seq(1968,2017,length=length(gin.correl))
tl2 = tl^2
polyGin = lm(gin.correl ~ tl + tl2)
summary(polyGin)
plot(gin.correl,lw=2,col="blue")
lines(polyGin$fit~tl,lw=2,col="red")

acf(gin.correl)
pacf(gin.correl)

```


```{r}
## Fit null ARIMA models (mean only) to each data set
nullGin = arima(gin.correl,order=c(0,0,0))

## Plot the null model predictions
#Remember null model is the one that takes the mean, as prediction
plot(gin.correl,xlim=c(1968,2030),lw=2,col="blue")
lines(predict(nullGin,n.ahead=150)$pred,lw=2,col="red")
lines(predict(nullGin,n.ahead=150)$pred+1.65*predict(nullGin,n.ahead=150)$se,lw=2,lty="dotted",col="red")
lines(predict(nullGin,n.ahead=150)$pred-1.65*predict(nullGin,n.ahead=150)$se,lw=2,lty="dotted",col="red")


ar1Gin = ar(gin.correl, order=1)
summary(ar1Gin)
ar1Gin

## Plot the ar model predictions
plot(gin.correl,xlim=c(1968,2030),lw=2,col="blue")
lines(predict(ar1Gin,n.ahead=150)$pred,lw=2,col="red")
lines(predict(ar1Gin,n.ahead=150)$pred+1.6*predict(ar1Gin,n.ahead=150)$se,lw=2,lty="dotted",col="red")
lines(predict(ar1Gin,n.ahead=150)$pred-1.5*predict(ar1Gin,n.ahead=150)$se,lw=2,lty="dotted",col="red")
```


```{r}

gin.spec <- spec.pgram(gin.correl,log="no", taper=0)

spGin <- gin.spec$freq*12

plot(gin.spec$spec~spGin,xlab="period",ylab="spectral density",type="h")
title("Gin periodogram")

spGin <- gin.spec$freq[order(gin.spec$spec)]
top3 <- tail(spGin,3)

top3

summary(spGin)
```

#Load image and find DFT
```{r}
library("EBImage")

image <- readImage("textDA.png")
image=EBImage::channel(image, "gray")
display(image, method="raster")

lincoln <- readImage("lincoln.jpg")
lincoln=EBImage::channel(lincoln, "gray")
display(lincoln, method="raster")

# This function is used to place place the dc-component in the middle (like a 2D fftshift)
# if m is a square 2^n matrix, center_matrix(center_matrix(m)) == m
center_matrix <- function(m) {
  dims <- dim(m)
  m1 <- cbind(m[,(dims[2]/2+1):dims[2]],m[,1:(dims[2]/2)])
  rbind(m1[(dims[1]/2+1):dims[1],],m1[1:(dims[1]/2),])
}

#Apply Fourier Transform
image_f <- fft(image)
lincoln_f <- fft(lincoln)

#compute magnitudes
image_m <- sqrt(abs(image_f))
lincoln_m <- sqrt(abs(lincoln_f))

#shift frequencies to the middle (like fftshift)
image_c <- center_matrix(image_m)
lincoln_c <- center_matrix(lincoln_m)

par(mfrow=c(1,2))
display(image, method="raster")
title("Original grayscale image")
display(image_c, method="raster")
title("2D DFT magnitude")

par(mfrow=c(1,2))
display(lincoln, method="raster")
title("Original grayscale image")
display(lincoln_c, method="raster")
title("2D DFT magnitude")
```

#Log of 2D DFT
log transform to visualise better the 2D DFT magnitude
```{r}
image.log <- 0.6*log(image_c, base = 10)
lincoln.log <- 0.6*log(lincoln_c, base = 10)

par(mfrow=c(1,2))
display(image,method="raster")
title("Original grayscale image")
display(image.log, method="raster")
title("Log transf of 2D DFT magnitude")

par(mfrow=c(1,2))
display(lincoln, method="raster")
title("Original grayscale image")
display(lincoln.log, method="raster")
title("Log transf of 2D DFT magnitude")
```
Both power spectrums show high frequency energy in waves along the X-direction and Y-direction although the text image is more clear so values are more visible all along the spectrum.

#low pass filter
```{r}
#Cut off threshold 55
sizeim =dim(image_f)[1]  #This is the size of the image (we are working with same width/height)
size=55       #Approx. size of the discoid object (we assume same width/height)
w0 = makeBrush(size = size, shape = 'disc')
#display(normalize(w0), method="raster")

# This creates a disc-sized element of 55, let's pad the rest with 0
# Note: There might be a more effective/different way to do this. Feel free to improve 
w1=padzeros(w0,(sizeim-size)/2,side="both")
w1=t(w1)
w=padzeros(w1,(sizeim-size)/2,side="both")
#that gives us a 255x255 filter, I need more zeros to pad
w=rbind(w,array(0,255))
w=cbind(w,array(0,256))
#Then I need to center my element such as that the discoid is in the center
w=center_matrix(center_matrix(w))

display(normalize(w), method="raster")

#Remember that filtering in frequency field means just a multiplication!
#Notice that we apply the filter on the DC-centered-image
clown_flow <- w * as.Image(center_matrix(image_f))
display(normalize(abs(clown_flow)), method="raster")

# place the DC component back to its place before the inverse fourier
clown_low  <- fft(center_matrix(clown_flow), inverse=TRUE)
display(normalize(abs(clown_low)), method="raster")


#Cut off threshold 105
sizeim =dim(image_f)[1]  #This is the size of the image (we are working with same width/height)
size=105       #Approx. size of the discoid object (we assume same width/height)
w0 = makeBrush(size = size, shape = 'disc')
#display(normalize(w0), method="raster")

# This creates a disc-sized element of 55, let's pad the rest with 0
# Note: There might be a more effective/different way to do this. Feel free to improve 
w1=padzeros(w0,(sizeim-size)/2,side="both")
w1=t(w1)
w=padzeros(w1,(sizeim-size)/2,side="both")
#that gives us a 255x255 filter, I need more zeros to pad
w=rbind(w,array(0,255))
w=cbind(w,array(0,256))
#Then I need to center my element such as that the discoid is in the center
w=center_matrix(center_matrix(w))

display(normalize(w), method="raster")

#Remember that filtering in frequency field means just a multiplication!
#Notice that we apply the filter on the DC-centered-image
clown_flow <- w * as.Image(center_matrix(image_f))
display(normalize(abs(clown_flow)), method="raster")

# place the DC component back to its place before the inverse fourier
clown_low  <- fft(center_matrix(clown_flow), inverse=TRUE)
display(normalize(abs(clown_low)), method="raster")

#Cut off threshold 200
sizeim =dim(image_f)[1]  #This is the size of the image (we are working with same width/height)
size=201     #Approx. size of the discoid object (we assume same width/height)
w0 = makeBrush(size = size, shape = 'disc')
#display(normalize(w0), method="raster")

# This creates a disc-sized element of 55, let's pad the rest with 0
# Note: There might be a more effective/different way to do this. Feel free to improve 
w1=padzeros(w0,(sizeim-size)/2,side="both")
w1=t(w1)
w=padzeros(w1,(sizeim-size)/2,side="both")
#that gives us a 255x255 filter, I need more zeros to pad
w=rbind(w,array(0,255))
w=cbind(w,array(0,256))
#Then I need to center my element such as that the discoid is in the center
w=center_matrix(center_matrix(w))

display(normalize(w), method="raster")

#Remember that filtering in frequency field means just a multiplication!
#Notice that we apply the filter on the DC-centered-image
clown_flow <- w * as.Image(center_matrix(image_f))
display(normalize(abs(clown_flow)), method="raster")

# place the DC component back to its place before the inverse fourier
clown_low  <- fft(center_matrix(clown_flow), inverse=TRUE)
display(normalize(abs(clown_low)), method="raster")

```
Thew low pass filter blurs the image and as the threshold increases the less powerful it is to blur. Therefore images with low threshold are more blury.

high pass filter
```{r}

sizeim = dim(image_f)[1]
size = 55
w0=makeBrush(size=size, shape='disc')
display(normalize(w0), method="raster")
  
# This creates a disc-sized element of 55, let's pad the rest with 0
# Note: There might be a more effective/different way to do this. Feel free to improve 
w1=padzeros(w0,((sizeim-size)/2),side="both")
w1=t(w1)
w=padzeros(w1,((sizeim-size)/2),side="both")
#that gives us a 255x255 filter, I need more zeros to pad
w=rbind(w,array(0,255))
w=cbind(w,array(0,256))
  
#Then I need to center my element such as that the discoid is in the center
w=center_matrix(center_matrix(w))
display(normalize(w), method="raster")
  
w <- apply(w, 1:2, function(x) x=abs(x-1))
  
#Remember that filtering in frequency field means just a multiplication!
#Notice that we apply the filter on the DC-centered-image
image_flow <- w * as.Image(center_matrix(image_f))
display(normalize(abs(image_flow)), method="raster")
  
# place the DC component back to its place before the inverse fourier
image_low  <- fft(center_matrix(image_flow), inverse=TRUE)
image_filtered <- normalize(abs(image_low))
display(image_filtered, method="raster")                            
```

The high pass filter brightens the edges in the images. Defining the threshold helps indicate how much attention we put to noise.

```{r}
artist1 <- readImage("artist1.jpg")
artist2 <- readImage("artist2.jpg")
artist1=EBImage::channel(artist1, "gray")
artist2=EBImage::channel(artist2, "gray")
display(artist1, method = "raster")
display(artist2, method = "raster")

artist1_f <- fft(artist1)
artist2_f <- fft(artist2)

sizeim = dim(artist1_f)[1]
size = 55
w0=makeBrush(size=size, shape='disc')

fill_in <- 1
if(mod(sizeim,2) == 0){
  fill_in <- 0
}

# This creates a disc-sized element of 55, let's pad the rest with 0
# Note: There might be a more effective/different way to do this. Feel free to improve 
w1=padzeros(w0,((sizeim-size)/2)-fill_in,side="both")
w1=t(w1)
w=padzeros(w1,((sizeim-size)/2)-fill_in,side="both")
#that gives us a 255x255 filter, I need more zeros to pad
w=rbind(w,array(0,sizeim-fill_in-1))
w=cbind(w,array(0,sizeim-fill_in))

#Then I need to center my element such as that the discoid is in the center
w=center_matrix(center_matrix(w))

#Remember that filtering in frequency field means just a multiplication!
#Notice that we apply the filter on the DC-centered-image
image_flow <- w * as.Image(center_matrix(artist1_f))
#display(normalize(abs(image_flow)), method="raster")

# place the DC component back to its place before the inverse fourier
image_low  <- fft(center_matrix(image_flow), inverse=TRUE)
art1.filt <- normalize(abs(image_low))

sizeim = dim(artist2_f)[1]
size = 55
w0=makeBrush(size=size, shape='disc')

fill_in <- 1
if(mod(sizeim,2) == 0){
  fill_in <- 0
}
# This creates a disc-sized element of 55, let's pad the rest with 0
# Note: There might be a more effective/different way to do this. Feel free to improve 
w1=padzeros(w0,((sizeim-size)/2)-fill_in,side="both")
w1=t(w1)
w=padzeros(w1,((sizeim-size)/2)-fill_in,side="both")
#that gives us a 255x255 filter, I need more zeros to pad
w=rbind(w,array(0,sizeim-fill_in-1))
w=cbind(w,array(0,sizeim-fill_in))

#Then I need to center my element such as that the discoid is in the center
w=center_matrix(center_matrix(w))

w <- apply(w, 1:2, function(x) x=abs(x-1))

#Remember that filtering in frequency field means just a multiplication!
#Notice that we apply the filter on the DC-centered-image
image_flow <- w * as.Image(center_matrix(artist2_f))
#display(normalize(abs(image_flow)), method="raster")

# place the DC component back to its place before the inverse fourier
image_low  <- fft(center_matrix(image_flow), inverse=TRUE)
art2.filt <- normalize(abs(image_low))
  
display(art1.filt+art2.filt, method = "raster")
title("Mixed images of artist with low and high pass filter")
```